VarCovar Rembrandt

### 2023-09-22 reporting/simu_rembrandt_breast.Rmd
update with type-I error

- Type (D) : $\beta \sim \mathcal{N}(0, 0.4^2)$
- Type (E) : half of genes follows $\beta \sim \mathcal{N}(-0.4, 0.2^2)$ and half follows $\beta \sim \mathcal{N}(0.4, 0.2^2)$
- Type (F) : half of genes follows $\beta \sim \mathcal{N}(-0.8, 0.4^2)$ and half follows $\beta \sim \mathcal{N}(0.8, 0.4^2)$
- Type (G) : $\beta \sim \mathcal{N}(0, 0.4^2)$. Correlation matrix of $\beta$ is the same as correlation matrix of genes.
- Type (H) : half of genes follows $\beta \sim \mathcal{N}(-0.4, 0.2^2)$ and half follows $\beta \sim \mathcal{N}(0.4, 0.2^2)$. Correlation matrix of $\beta$ is the same as correlation matrix of genes.
- Type (I) : half of genes follows $\beta \sim \mathcal{N}(-0.8, 0.4^2)$ and half follows $\beta \sim \mathcal{N}(0.8, 0.4^2)$. Correlation matrix of $\beta$ is the same as correlation matrix of genes.

@@

@@

- Global Test is fast compared to other approach. It is not completely clear to me why because it also relies on permutation and is similar to Wald test. There is probably a clever implementation ?
- In terms of power Wald test is the best option, closely followed by Global test
- sGBJ is slow when the number of genes rise. This is due to permutation for variance matrix estimation that might be reduced to improve computation time.
- sGBJ performance is similar to most approach except when the number of genes increases.

## Comprehend tests

@@

### Wald test

@@

### Global test

@@

### Global boost test

@@

### Thinking

>>>>

- Type (Z) : $\beta = 0$ evaluates the type-I error
- Type (D) : $\beta \sim \mathcal{N}(0, 0.4^2)$
- Type (E) : half of genes follows $\beta \sim \mathcal{N}(-0.4, 0.2^2)$ and half follows $\beta \sim \mathcal{N}(0.4, 0.2^2)$
- Type (F) : half of genes follows $\beta \sim \mathcal{N}(-0.8, 0.4^2)$ and half follows $\beta \sim \mathcal{N}(0.8, 0.4^2)$
- Type (G) : $\beta \sim \mathcal{N}(0, 0.4^2)$. Correlation matrix of $\beta$ is the same as correlation matrix of genes.
- Type (H) : half of genes follows $\beta \sim \mathcal{N}(-0.4, 0.2^2)$ and half follows $\beta \sim \mathcal{N}(0.4, 0.2^2)$. Correlation matrix of $\beta$ is the same as correlation matrix of genes.
- Type (I) : half of genes follows $\beta \sim \mathcal{N}(-0.8, 0.4^2)$ and half follows $\beta \sim \mathcal{N}(0.8, 0.4^2)$. Correlation matrix of $\beta$ is the same as correlation matrix of genes.

@@

```{r fig.height=6, fig.cap="Type-I error (Type = Z). N is the number of patients and NG the number of genes in the pathway."}
plot_simu_alpha
```

@@

- Global Test is fast compared to other approach. It is not completely clear to me why because it also relies on permutation and is similar to Wald test. There is probably a clever implementation ?
- In terms of power Wald test is the best option, closely followed by Global test
- Type-I error is well controlled by the different methods. We note that sGBJ is too conservative when the number of genes is high.
- sGBJ is slow when the number of genes rise. This is due to permutation for variance matrix estimation that might be reduced to improve computation time.
- sGBJ performance is similar to most approach except when the number of genes increases which is partly explained by the over-conservative control of Type-I error.

# Notes

@@

## Wald test

@@

## Global test

@@

## Global boost test

@@

## Thinking
